% This document is part of the transientdict project.
% Copyright 2013 the authors.

\documentclass[12pt]{article}

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\Fermi}{\project{Fermi}}
\newcommand{\given}{\,|\,}
\newcommand{\dd}{\mathrm{d}}
\renewcommand{\count}{y}
\newcommand{\pars}{\theta}
\newcommand{\mean}{\lambda}
\newcommand{\Poisson}{{\mathcal P}}
\newcommand{\Uniform}{{\mathcal U}}
\newcommand{\bg}{\mathrm{bg}}
\newcommand{\word}{\phi}

\begin{document}

\section*{A model for magnetar bursts}

\noindent
Huppenkothen, Murray, Hogg, Brewer, Frean \\
\textsl{2013 December}

\paragraph{abstract}
Magnetars are whatever and people need whatever.
Magnetars produce bursts,
  each of which appears to be composed of a set of one or more spikes.
Here we build a probabilistic generative model for the \Fermi\ GBM photon data on Magnetar bursts.
The spikes appear to be somewhat asymmetric
  (rise looks different from decay),
  have various amplitudes,
  but have widths that appear to be similar \emph{within} each burst.
We model each burst as being composed of a mixture of spikes,
  where each spike is a scaled, stretched, and shifted version of a universal dimensionless function.
We perform probabilistic inference with a Poisson likelihood and vague priors to determine
  the positions, amplitudes, and positions of the spikes,
  the width of the spikes in each burst,
  and the shape of the dimensionless spike function.
The phenomenology of spike multiplicity, shape, and width is discussed.

\section{introduction}

Magnetars make bursts.
Magnetar bursts have many spikes each.
The spikes look similar within each burst.
That motivates a very simple model,
  in which each burst is made up of a sum of spike models, which we call words.

\section{data}

The ``raw'' data, for our purposes, are photon arrival times.
These come from triggered photon streams telemetered down from the \Fermi\ Gamma-ray Burst Monitor (GBM).
These data streams have [the following crazy properties, including dead time, saturation]

There are many small details.
We choose [what binning and why?].
We use data [in what interval?] around each burst.

In what follows, the data for one burst are photon counts $\count_n$ (integer) in $N$ bins $n$.
The data from a few example bursts are shown in Figure [what figure?].

\section{model}

The model for the data in or near one burst is
\begin{eqnarray}
p(\count_n\given\pars) &=& \Poisson(\count_n\given\mean_n)
\\
\mean_n &=& \mean_{\bg} + \sum_{k=1}^K \mean_{nk}
\\
\mean_{nk} &\equiv& \int_{t_n-\Delta/2}^{t_n+\Delta/2} A_k\,\word(\frac{t'-t_k}{\tau})\,\dd t'
\\
\word(\xi) &=& \left\{\begin{array}{ll}\exp(\xi) & \mbox{for $\xi<0$}\\ \exp(-\xi/s) & \mbox{for $\xi\geq 0$}\end{array}\right.
\\
\theta &\equiv& [s, \mean_{\bg}, \tau, \{A_k, t_k\}_{k=1}^K ]
\quad,
\end{eqnarray}
where $\Poisson(\count\given\mean)$ is the Poisson probability of getting count $y$ given mean rate $\mean$,
  $\pars$ is the vector or blob of all model parameters,
  $\mean_{\bg}$ is the background (DC) level in the bin,
  $t_n$ is the time of the center of the bin,
  $\Delta$ is the full width of the bin,
  $K$ is the number of words (spikes) $k$ making up the burst,
  $\phi(\xi)$ is the dimensionless word function,
  $A_k, t_k$ are the amplitudes and time offsets of the words,
  and $s, \tau$ set the shape (asymmetry or skew) and duration of the words.
Figure [what figure?] shows the shape of a word
  and an example of a burst generated by our model.
[This figure should show a word in $\xi$ space,
  a multi-word burst in $t$ space,
  and a Poisson draw from that model.]

The model for a large set of bursts is whatever.
[We should probably, in the above, index bursts $m$,
  and maybe even magnetar $j$.]

\emph{Notes to selves:}
At one point we figured we could \emph{fix} shape parameter $s$
  to be the same for all the words in all the bursts from one magnetar,
  but permit the duration $\tau$ to vary.
However, as we worked in our coding session we moved more in the direction
  of thinking in terms of ``rise time'' $\tau$ and decay time $s\,\tau$.
It might be better to parameterize that way.
One possible point of discovery would be that the rise times
  are fundamental to each magnetar,
  but decay times vary.
(For instance.)

The prior pdfs are
\begin{eqnarray}
p(\ln\lambda_{\bg}) &=& \Uniform(\ln\lambda_{\bg}\given a_1, a_2)
\\
p(\ln s) &=& \Uniform(\ln s\given a_3, a_4)
\\
p(\ln\tau) &=& \Uniform(\ln\tau\given a_5, a_6)
\\
p(\ln A_k) &=& \Uniform(\ln A_k\given a_7, a_8)
\\
p(t_k) &=& \Uniform(t_k\given a_9, a_{10})
\quad,
\end{eqnarray}
where $\Uniform(x\given a, b)$ is the uniform distribution for $x$ in the range $a<x<b$.

We use \project{emcee} (CITE) to perform MCMC sampling.
[We initialize the MCMC walkers how?]

Finally, there is the issue of how to set the number $K$ of words in each burst.
We take a ``cross-validation'' approach [take that, BJB],
  in which we ask what value of $K$ does the best job at making posterior predictions
  about left-out data.
[We construct this validation test how?]
Figure [what figure?] shows an example burst,
  fit with $K=1$, $2$, $3$, $4$, and $5$ words.
It also shows the result of the posterior-prediction validation
  that selects $K=4$ as the best choice for this burst.

\section{results}

\section{discussion}

\paragraph{acknowledgements}
We thank the organizers of MaxEnt2013.

\end{document}
